---
title: "EDA-challenge"
format: html
editor: visual
---

Exercise 3 
Steps 1-3
```{r}
library(tidyverse)
library(ggplot2)
f<-"https://raw.githubusercontent.com/difiore/ada-datasets/main/data-wrangling.csv" #choose my dataset
d<- read_csv(f, col_names = TRUE)
names(d) #view my variable names
d<- d %>%
  mutate(BSD = Body_mass_male_mean/Body_mass_female_mean,
         sex_ratio = AdultFemale/AdultMales,
         diameter = 2 * sqrt(HomeRange_km2/pi),
         DI = DayLength_km/diameter)
view(d) #checking my new variables
```


Plotting
Step 4
```{r}
#overall plot
plot1<- ggplot(data = d, aes(x = Move, y = DayLength_km, color = Family)) +
  geom_point(na.rm = TRUE) +
  xlab("Time Spent Moving") +
  ylab("Day Range Length (km)")
plot1
#By family using facet wrap
plot2<- d %>%
  filter(!is.na(Move), !is.na(DayLength_km)) %>% #prevents family's with no data in these variables from facet wrapping into my plot
  ggplot(aes(x = Move, y = DayLength_km, color = Family)) +
  geom_point(na.rm = TRUE) +
  xlab("Time Spent Moving") +
  ylab("Day Range Length (km)") +
  facet_wrap(vars(Family))
plot2
```
Step 4 Questions:
Do species that spend more time moving travel farther overall? No. Overall it appears that regardless of time spent moving, most species travel within 0-3 km in a day.

How about within any particular primate family? Cercopithecidae traveled the furthest and spent the most time moving. However, the individuals that moved the most were not the ones that traveled the furthest in a day.

Should you transform either of these variables? Why or why not? I don't think you need to transform the data as neither appear skewed. When doing a log transformation of the variables, it led to negative values in Day Range Length which is something we would avoid doing.








